---
title: "Grando-8"
author: "John Grando"
date: "April 21, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

## Grando 8 Lab

```{r}
if (Sys.info()["sysname"]=="Windows"){
  setwd("~/Masters/DATA606/Week8/Lab/Lab8")
} else {
  setwd("~/Documents/Masters/DATA606/Week8/Lab/Lab8")
}
load("more/evals.RData")
require(ggplot2)
```

####Exercise 1 - Is this an observational study or an experiment? The original research question posed in the paper is whether beauty leads directly to the differences in course evaluations. Given the study design, is it possible to answer this question as it is phrased? If not, rephrase the question.

Answer:

This is an observational study.  No, since this is an observational study it is only generally sufficient to show associations between variables and should not be used to draw causal conclusions.  The appropriate way to rephrase the question would be to ask: Is there a correlation between the instructor's beauty and evaluation scores?


####Exercise 2 - Describe the distribution of score. Is the distribution skewed? What does that tell you about how students rate courses? Is this what you expected to see? Why, or why not?

Answer:

```{r}
ggplot(evals, aes(x=score))+geom_histogram(binwidth = 0.125, position = "identity", aes(y=..density..))+stat_function(fun=dnorm,color="black",args=list(mean=mean(area_df$data),sd(area_df$data)))+labs(x="Score")
qqnorm(evals$score)
qqline(evals$score)
```

The distribution has a left skew.  Additionally, we can see that the residuals have a steped nature which show that there are relatively few possible unique values that were recorded.  This shows that students will typically give a high score to an average class since the median is much closer to the maximum value than the minimum.  It leads me to assume that if a student felt neutral towards a teacher's performance, they generally give them a good score, possibly giving the professor the benefit of the doubt about their skills.  I did expect to see something similar to this.  In my experience, I find that people generally will lean towards being aggreeable in these type of surveys since it doesn't require an argument and at least, in their minds, they would not have to strongly defend a higher score rather than a lower.

####Exercise 3 - Excluding score, select two other variables and describe their relationship using an appropriate visualization (scatterplot, side-by-side boxplots, or mosaic plot).

Answer:

for this exercise, I will compare the percent of students who completed the evaluations against the average beauty of the professo.

```{r}
ggplot(evals, aes(x=cls_perc_eval, y=bty_avg)) + geom_point()+ geom_smooth(method = lm, fullrange = TRUE)
e3.lm <- lm(bty_avg ~ cls_perc_eval, data = evals)
summary(e3.lm)
```

From the linear regression line, there appears to be a slightly positive slope between the percent of students who completed the evaluation and the average beauty rating of the professor.  However, it does not appear that there is a good r-squared correlation with a linear model and there are not any apparent patterns for a better fit in other models (e.g. quadratic, power, etc.).  It is noted that there is a significantly low p-value for the slope of the fit which suggests that the class percentage may be a statistically significant predictor of average beauty even though the correlation is low.

####Exercise 4 - Replot the scatterplot, but this time use the function jitter() on the y- or the x-coordinate. (Use ?jitter to learn more.) What was misleading about the initial scatterplot?

Answer:

```{r}
plot(evals$score ~ evals$bty_avg)
jitter_bty_avg <- jitter(evals$bty_avg, factor = 1.5)
plot(evals$score ~ jitter_bty_avg)
```

Due to the low number of unique variable values possible, there appear to be many overlapping data points.  The jitter plot slightly shifts the vector values so that this can be more apparent.

####Exercise 5 - Let’s see if the apparent trend in the plot is something more than natural variation. Fit a linear model called m_bty to predict average professor score by average beauty rating and add the line to your plot using abline(m_bty). Write out the equation for the linear model and interpret the slope. Is average beauty score a statistically significant predictor? Does it appear to be a practically significant predictor?

Answer:

```{r}
m_bty <- lm(score ~ bty_avg, data = evals)
summary(m_bty)
plot(evals$score ~ jitter_bty_avg)
abline(m_bty)
```

$$evaluation\quad score\quad =\quad 3.88\quad +\quad 0.067\quad *\quad average\quad beauty$$

For every one point of increase in average beauty, a professor is expected to recieve an additiona 0.067 points in their evaluation.  Since the p-value is less than 0.05, it appears that average beauty is a statistically significant predictor of the evaluation score.  It does not appear to be a practically significant predictor because the r-squared value is 0.03 which means that only 3% of the variability is explained by the linear model.  Additionally, the slope of the equation is very small and may not have a significant impact on the results.

####Exercise 6 - Use residual plots to evaluate whether the conditions of least squares regression are reasonable. Provide plots and comments for each one (see the Simple Regression Lab for a reminder of how to make these).

Answer:

```{r}
hist(m_bty$residuals)
qqnorm(m_bty$residuals)
qqline(m_bty$residuals)

residuals <- resid(m_bty)
residuals <- data.frame(cbind(evals$bty_avg, residuals))
names(residuals) <- c("avg_bty", "residuals")
ggplot(residuals, aes(y=residuals, x=avg_bty))+geom_point()+geom_hline(yintercept = 0)
```


Give an analysis of each plot... check the book

####Exercise 7 - P-values and parameter estimates should only be trusted if the conditions for the regression are reasonable. Verify that the conditions for this model are reasonable using diagnostic plots.

Answer:



####Exercise 8 - Is bty_avg still a significant predictor of score? Has the addition of gender to the model changed the parameter estimate for bty_avg?

Answer:



####Exercise 9 - What is the equation of the line corresponding to males? (Hint: For males, the parameter estimate is multiplied by 1.) For two professors who received the same beauty rating, which gender tends to have the higher course evaluation score?

Answer:



####Exercise 10 - Create a new model called m_bty_rank with gender removed and rank added in. How does R appear to handle categorical variables that have more than two levels? Note that the rank variable has three levels: teaching,  tenure track, tenured.

Answer:



####Exercise 11 - Which variable would you expect to have the highest p-value in this model? Why? Hint: Think about which variable would you expect to not have any association with the professor score.

Answer:



####Exercise 12 - Check your suspicions from the previous exercise. Include the model output in your response.

Answer:



####Exercise 13 - Interpret the coefficient associated with the ethnicity variable.

Answer:



####Exercise 14 - Drop the variable with the highest p-value and re-fit the model. Did the coefficients and significance of the other explanatory variables change? (One of the things that makes multiple regression interesting is that coefficient estimates depend on the other variables that are included in the model.) If not, what does this say about whether or not the dropped variable was collinear with the other explanatory variables?

Answer:



####Exercise 15 - Using backward-selection and p-value as the selection criterion, determine the best model. You do not need to show all steps in your answer, just the output for the final model. Also, write out the linear model for predicting score based on the final model you settle on.

Answer:



####Exercise 16 - Verify that the conditions for this model are reasonable using diagnostic plots.

Answer:



####Exercise 17 - The original paper describes how these data were gathered by taking a sample of professors from the University of Texas at Austin and including all courses that they have taught. Considering that each row represents a course, could this new information have an impact on any of the conditions of linear regression?

Answer:



####Exercise 18 - Based on your final model, describe the characteristics of a professor and course at University of Texas at Austin that would be associated with a high evaluation score.

Answer:



####Exercise 19 - Would you be comfortable generalizing your conclusions to apply to professors generally (at any university)? Why or why not?

Answer: