---
title: "Grando-Final"
author: "John Grando"
date: "May 18, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=90),tidy=TRUE)
```

## Grando - Final

###Part I

####Question 1 - A student is gathering data on the driving experiences of other college students. A description of the data car color is presented below. Which of the variables are quantitative and discrete?

Answer:

d. daysDrive, gasMonth


daysDrive- Since this is indicated to be a count of days then it appears this value can only take on whole non-negative integers for the number of days the student drives.

gasMonth - This value is indicated to be cost, which is only calculated down to $0.01.  Therefore, there are not an infinite number of possible values for a given variable range.  Also, not that if this value were calculated in cents, then it would be all non-negative integer which is similar to the daysDrive variable.

####Question 2 - A histogram of the GPA of 132 students from this course in Fall 2012 class is presented below. Which estimates of the mean and median are most plausible?

Answer:

Since the data appears to have a left skew, we can immediately remove any option that states the mean is greater than the median (b, d).

when we evaluate the histogram we can see that if we add the percentageS right to left, approximately 41% of the data is above 3.5; therefore the median must be 3.5 or below.  This leaves us with one option which is:

a. mean = 3.3, median = 3.5

####Question 3 - A researcher wants to determine if a new treatment is effective for reducing Ebola related fever. What type of study should be conducted in order to establish that the treatment does indeed cause improvement in Ebola patients?

Answer:

a. Randomly assign Ebola patients to one of two groups, either the treatment or placebo group, and then compare the fever of the two groups

Option d (Both a and b) is not acceptable because it is not guaranteed that the sample is randomized and could contain an unknown bias.  Additionally, by not having a placebo group, we cannot be certain that we have appropriately set up the experiment to compare for a difference of only one variable (i.e. the treament) 

####Question 4 - A study is designed to test whether there is a relationship between natural hair color (brunette, blond, red) and eye color (blue, green, brown). If a large χ2 test statistic is obtained, this suggests that:

Answer:

c. there is an association between natural hair color and eye color.

####Question 5 - A researcher studying how monkeys remember is interested in examining the distribution of the score on a standard memory task. The researcher wants to produce a boxplot to examine this distribution. Below are summary statistics from the memory task. What values should the researcher use to determine if a particular score is a potential outlier in the boxplot?

Answer:

To determine possible outliers, the researcher would examine points that fell beyond the whisker limits of the boxplot, which are 1.5 * IQR.

```{r}
Q1 <- 37 
Q3 <- 49.8
IQR <- Q3 - Q1
(Lower_whisker <- 37 - 1.5 * IQR)
(Upper_whisker <- 49.8 + 1.5 * IQR)
```

b. 17.8 and 69.0

####Question 6 - The [blank] are resistant to outliers, whereas the [blank] are not.

Answer:

d. median and interquartile range; mean and standard deviation

####Question 7 - Figure A below represents the distribution of an observed variable. Figure B below represents the distribution of the mean from 500 random samples of size 30 from A. The mean of A is 5.05 and the mean of B is 5.04. The standard deviations of A and B are 3.22 and 0.58, respectively.

####a. Describe the two distributions (2 pts).

Answer:



####b. Explain why the means of these two distributions are similar but the standard deviations are not (2 pts).

Answer:



####c. What is the statistical principal that describes this phenomenon (2 pts)?

Answer:



###Part II

Consider the four datasets, each with two columns (x and y), provided below.

```{r}
#options(digits=2)
data1 <- data.frame(x=c(10,8,13,9,11,14,6,4,12,7,5),
y=c(8.04,6.95,7.58,8.81,8.33,9.96,7.24,4.26,10.84,4.82,5.68))
data2 <- data.frame(x=c(10,8,13,9,11,14,6,4,12,7,5),
y=c(9.14,8.14,8.74,8.77,9.26,8.1,6.13,3.1,9.13,7.26,4.74))
data3 <- data.frame(x=c(10,8,13,9,11,14,6,4,12,7,5),
y=c(7.46,6.77,12.74,7.11,7.81,8.84,6.08,5.39,8.15,6.42,5.73))
data4 <- data.frame(x=c(8,8,8,8,8,8,8,19,8,8,8),
y=c(6.58,5.76,7.71,8.84,8.47,7.04,5.25,12.5,5.56,7.91,6.89))

```

For each column, calculate (to two decimal places):

####a. The mean (for x and y separately; 1 pt).

Answer:

```{r}
#### Always show to decimal places?
options(digits = 2)
(data1_mean_x <- mean(data1$x))
(data1_mean_y <- mean(data1$y))

(data2_mean_x <- mean(data2$x))
(data2_mean_y <- mean(data2$y))

(data3_mean_x <- mean(data3$x))
(data3_mean_y <- mean(data3$y))

(data4_mean_x <- mean(data4$x))
(data4_mean_y <- mean(data4$y))
```


####b. The median (for x and y separately; 1 pt).

Answer:

```{r}
#### Always show to decimal places?
options(digits = 2)
(data1_median_x <- median(data1$x))
(data1_median_y <- median(data1$y))

(data2_median_x <- median(data2$x))
(data2_median_y <- median(data2$y))

(data3_median_x <- median(data3$x))
(data3_median_y <- median(data3$y))

(data4_median_x <- median(data4$x))
(data4_median_y <- median(data4$y))
```


####c. The standard deviation (for x and y separately; 1 pt).

Answer:


```{r}
#### Always show to decimal places?
options(digits = 2)
(data1_sd_x <- sd(data1$x))
(data1_sd_y <- sd(data1$y))

(data2_sd_x <- sd(data2$x))
(data2_sd_y <- sd(data2$y))

(data3_sd_x <- sd(data3$x))
(data3_sd_y <- sd(data3$y))

(data4_sd_x <- sd(data4$x))
(data4_sd_y <- sd(data4$y))
```


For each x and y pair, calculate (also to two decimal places; 1 pt):

####d. The correlation (1 pt).

Answer:

```{r}
(cor(data1$x, data1$y))
(cor(data2$x, data2$y))
(cor(data3$x, data3$y))
(cor(data4$x, data4$y))
```


####e. Linear regression equation (2 pts).

Answer:

```{r}
lm_d1 <- lm(y ~ x, data = data1)
summary(lm_d1)
lm_d2 <- lm(y ~ x, data = data2)
summary(lm_d2)
lm_d3 <- lm(y ~ x, data = data3)
summary(lm_d3)
lm_d4 <- lm(y ~ x, data = data4)
summary(lm_d4)
```


####f. R-Squared (2 pts).

Answer:

```{r}
cor(data1$y, data1$x)^2
summary(lm(y ~ x, data = data1))[[8]]

cor(data2$y, data2$x)^2
summary(lm(y ~ x, data = data2))[[8]]

cor(data3$y, data3$x)^2
summary(lm(y ~ x, data = data3))[[8]]

cor(data4$y, data4$x)^2
summary(lm(y ~ x, data = data3))[[8]]
```


####For each pair, is it appropriate to estimate a linear regression model? Why or why not? Be specific as to why for each pair and include appropriate plots! (4 pts)

Answer:

1. First we should visually inspect the data to ensure it has a linear pattern.

```{r}
require(ggplot2)
ggplot(data1, aes(y=y, x=x))+geom_point()+ geom_smooth(method = lm, fullrange = TRUE)
ggplot(data2, aes(y=y, x=x))+geom_point()+ geom_smooth(method = lm, fullrange = TRUE)
ggplot(data3, aes(y=y, x=x))+geom_point()+ geom_smooth(method = lm, fullrange = TRUE)
ggplot(data4, aes(y=y, x=x))+geom_point()+ geom_smooth(method = lm, fullrange = TRUE)
```

2. the residuals of the model are nearly normal

```{r}
#data1
ggplot(data = lm_d1, aes(x=resid(lm_d1))) + geom_histogram(binwidth = 1, position = "identity", aes(y=..density..)) + stat_function(fun=dnorm,color="black",args=list(mean=mean(lm_d1$residuals),sd(lm_d1$residuals))) + labs(x="Residuals")
#data2
ggplot(data = lm_d2, aes(x=resid(lm_d2))) + geom_histogram(binwidth = 1, position = "identity", aes(y=..density..)) + stat_function(fun=dnorm,color="black",args=list(mean=mean(lm_d2$residuals),sd(lm_d2$residuals))) + labs(x="Residuals")
#data3
ggplot(data = lm_d3, aes(x=resid(lm_d3))) + geom_histogram(binwidth = 1, position = "identity", aes(y=..density..)) + stat_function(fun=dnorm,color="black",args=list(mean=mean(lm_d3$residuals),sd(lm_d3$residuals))) + labs(x="Residuals")
#data3
ggplot(data = lm_d4, aes(x=resid(lm_d4))) + geom_histogram(binwidth = 1, position = "identity", aes(y=..density..)) + stat_function(fun=dnorm,color="black",args=list(mean=mean(lm_d4$residuals),sd(lm_d4$residuals))) + labs(x="Residuals")
```


3. the variability of the residuals is nearly constant



4. the residuals are independent



####Explain why it is important to include appropriate visualizations when analyzing data. Include any visualization(s) you create. (2 pts)

Answer:



